---
output:
  github_document:
    html_preview: false
---

```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/REAsDME-",
  out.width = "100%"
)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

# sfarrow: An R implementation to read/write  `sf` spatial objects from/to Parquet files

`sfarrow` is an experimental and work-in-progress package for testing how to
read/write Parquet files with `sf` objects in `R`.

Simple features are a popular format for representing spatial vector data using
`data.frames` and a list-like geometry column, implemented in the R package
[`sf`](https://r-spatial.github.io/sf/). Parquet files are an open-source,
column-oriented data storage format from Apache
([https://parquet.apache.org/](https://parquet.apache.org/)) which enable
efficient read/writing for large files. Parquet files are becoming popular
across programming languages and can be used in `R` using the package
[`arrow`](https://github.com/apache/arrow/).

The `sfarrow` implementation translates simple feature data objects using
well-known binary (WKB) format for geometry and reads/writes Parquet files. A
key goal of the package is for interoperability of the Parquet files
(particularly with Python `GeoPandas`), so coordinate reference system
information is maintained in a standard metadata format 
([https://github.com/geopandas/geo-arrow-spec](https://github.com/geopandas/geo-arrow-spec)).
Note to users: this metadata format is not yet stable for production uses and
may change in the future.

## Installation

Installation of the `sfarrow` package is through Github:

```{r eval=FALSE}
devtools::install_github("wcjochem/sfarrow@main")
```

Load the library to begin using
```{r}
library(sfarrow)
```

### `arrow` package

The installation requires the Arrow library which should be installed with the
`R` package `arrow` dependency. However, some systems may need to follow
additional steps to enable full support of that library. Please refer to the
`arrow`
[documentation](https://cran.r-project.org/web/packages/arrow/vignettes/install.html).

## Basic usage

Reading Parquet data of spatial files created with Python `GeoPandas`.
```{r}
# load Natural Earth low-res dataset. 
# Created in Python with geopandas.to_parquet()
path <- system.file("extdata", "world.parquet", package = "sfarrow")

world <- st_read_parquet(path)

head(world)
plot(sf::st_geometry(world))
```

Writing `sf` objects to Parquet format files. These Parquet files created with
`sfarrow` can be read within Python using `GeoPandas`.
```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package="sf"), quiet=TRUE)

st_write_parquet(obj=nc, dsn=file.path(tempdir(), "nc.parquet"))

# read back into R
nc_p <- st_read_parquet(file.path(tempdir(), "nc.parquet"))

head(nc_p)
plot(sf::st_geometry(nc_p))
```

For additional examples please see the vignettes.

## Contributions
Contributions, questions, ideas, and issue reports are welcome. Please raise an
issue to discuss or submit a pull request.

## Acknowledgements
This work benefited from the work by developers in the GeoPandas, Arrow, and
r-spatial teams. Thank you to the teams for their excellent, open-source work.

