---
output:
  github_document:
    html_preview: false
---

```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/REAsDME-",
  out.width = "100%"
)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

# sfarrow: An R implementation to read/write between Simple Feature (`sf`)
spatial objects to parquet files.

`sfarrow` is an experimental and work-in-progress file for testing how to
read/write Parquet files from `sf` objects.

Simple features are a popular format for representing spatial vector data using
`data.frames` and a list-like geometry column, implemented in the R package
`[sf](https://r-spatial.github.io/sf/)`. Parquet files are an open-source,
column-oriented data storage formate from Apache
([https://parquet.apache.org/](https://parquet.apache.org/)) which enable
efficient read/writing for large files. Parquet files are also becoming popular
across programming languages and can be used in `R` using the package
`[arrow](https://github.com/apache/arrow/)`. The `sfarrow` implementation
translates simple feature data objects and reads/writes Parquet files. A key
goal is for interoperability of the parquet files (particularly with Python
`GeoPandas`), so coordinate reference system information is maintained in a
standard metadata format
([https://github.com/geopandas/geo-arrow-spec](https://github.com/geopandas/geo-arrow-spec)).

## Installation
Installation of the `sfarrow` package is through Github:

```{r}
devtools::install_github("https://github.com/wcjochem/sfarrow@main")
```

## Basic usage

Reading Parquet data of spatial files created with `GeoPandas`.
```{r}
# load Natural Earth low-res dataset. 
# Created from GeoPandas.to_parquet()
path <- system.file("extdata", "world.parquet", package = "sfarrow")

world <- st_read_parquet(path)

head(world)
plot(sf::st_geometry(world))
```

Writing `sf` objects to Parquet format files.
```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))

st_write_parquet(obj=nc, dsn=file.path(tempdir(), "nc.parquet"))

# read back into R
nc_p <- st_read_parquet(file.path(tempdir(), "nc.parquet"))

head(nc_p)
plot(sf::st_geometry(nc_p))
```

These Parquet files created with `sfarrow` can be read within Python using `GeoPandas`.
```{python, eval=FALSE, echo=TRUE, python.reticulate=FALSE}
nc = geopandas.read_parquet('path/to/file/nc.parquet')
```

## Acknowledgements

```{r}
citation("sfarrow")
```
This work benefited from the work by developers in the GeoPandas, Arrow, and
r-spatial teams. Thank you to their excellent, open-source work.
